---
title: "FISH 421: Lab 1: Logistic population growth"
author: "Milo Adkison"
date: "April 19, 2016"
output: html_document
---
Step 1. Use the logistic growth equation to calculate abundance by year, start with 1 individual over a period of 100 years 
Intrinsic population growth rate at low abundance

```{r, eval = F}
r = 0.2
```

Carrying capacity

```{r, eval = F}
K = 1000
```


Create a function to calculate abundance, the inputs are *r*, *K*, *N*, and *t*.


```{r}
logistic_growth <- function(r, K, N, t){  #create a function with the input values
   z =  NA                                #define a vector for holding the calculated values
   z[1] = N                               # define the starting value
   for(i in 1:(t-1)){                     #run iterations to one less than the total time (you've provided the first value)
      z[i+1] = z[i]+ r * z[i] * (1-z[i]/K)#logistic growth model
   }
   df <- data.frame(year = 1:t, N = z, K = factor(K), r = factor(r))   #create a dataframe for holding the values
}
```

run the model with selected inputs - name it so you can plot later

```{r}
dat <- logistic_growth(0.2, 1000, 1, 100) 
```

Graph the abundance over time
```{r}
library(ggplot2)
ggplot(dat, aes(year, N))+geom_line()
```

Change the initial abundance (in year 1) to 1500. What does the population do?
```{r}
dat1 <- logistic_growth(0.2, 1000, 1500, 100) #give the model run a different name to be able to compare the results
```

Plot the results of both iterations
```{r, message = F}
library(plyr)
library(dplyr)
dat %>% 
   mutate(model = 'one') -> dat

dat1 %>% 
      mutate(model = 'two') -> dat1

dat2 <- rbind(dat,dat1)

ggplot(dat2, aes(year, N, color=model))+geom_line()
```

Change logistic growth model to calculate what abundance the population would have if it started from the value in vector *N(t)* the year before. Add a replacement line (red).
```{r}
dat3 <- data.frame(N = seq(0,1500,by=15), r = 0.2, K = 1000) #create a data.frame of abundance in increments of 15
dat3 %>%
mutate(N1 = N + r * N * (1-N/K)) %>% 
ggplot(., aes(N, N1))+geom_point()+geom_line(aes(N,N), color='red')
```


Change the parameters. What happens if you change K
```{r, warning = F}
k100 <- logistic_growth(0.2, 100, 1, 100)
k500 <- logistic_growth(0.2, 500, 1, 100)
k1000 <- logistic_growth(0.2, 1000, 1, 100)
k1500 <- logistic_growth(0.2, 1500, 1, 100)
k2000 <- logistic_growth(0.2, 2000, 1, 100)
k5000 <- logistic_growth(0.2, 5000, 1, 100)

rbind_all(list(k100 ,k500, k1000, k1500, k2000, k5000)) %>% 
      ggplot(., aes(year, N, color=K))+geom_line()
```

What do you think will happen if we double ‘r’ from 0.2 to 0.4? 
Think about this for a few minutes before you try it.  

Return *K* to 1000

```{r}
r20 <- logistic_growth(0.2, 1000, 1, 100)
r40 <- logistic_growth(0.4, 1000, 1, 100)

r2040 <- rbind(r20, r40) 

r2040 %>% 
   ggplot(aes(year, N, color=r))+geom_line()
```

What changed? What didn’t change? Why? 

Look at the graph of *N(t+1)* vs. *N(t)*. How does it change when *r* is changed 
from 0.2 to 0.4. What’s the change in the amount of density-dependence?
```{r}
dat4 <- data.frame(N = seq(0,1500,by=15), r2 = 0.2, r4 = 0.4, K= 1000) #create a data.frame of abundance in increments of 15

dat4 %>%
mutate(N2 = N + r2 * N * (1-N/K),
       N4 = N + r4 * N * (1-N/K)) %>% 
ggplot(., aes(N, N2))+geom_line()+geom_line(aes(N,N), color='red', lty = 4)+
   geom_line(aes(N, N4), color='blue')
```

Try it with *r* = 1
```{r}
dat5 <- data.frame(N = seq(0,1500,by=15), r2 = 0.2, r4 = 0.4, r10 = 1, K= 1000) #create a data.frame of abundance in increments of 15

dat5 %>%
mutate(N2 = N + r2 * N * (1-N/K),
       N4 = N + r4 * N * (1-N/K),
       N10 =N + r10 * N * (1-N/K) ) %>% 
ggplot(., aes(N, N2))+geom_line()+geom_line(aes(N,N), color='red', lty = 4)+
   geom_line(aes(N, N4), color='blue')+
   geom_line(aes(N, N10), color='green')
```

Abundance vs time
```{r}
r100 <- logistic_growth(1, 1000, 1, 100)

r20401 <- rbind(r2040, r100) 

r20401 %>% 
   ggplot(aes(year, N, color=r))+geom_line()
```

Try *r* = 2.0. What’s that funny little squiggle there starting about year 15? Look at the abundance for years 8-20. Is the population climbing steadily to its carrying capacity? What is it doing?

```{r}
dat6 <- data.frame(N = seq(0,1500,by=15), r2 = 0.2, r4 = 0.4, r10 = 1,r20 = 2, K= 1000) #create a data.frame of abundance in increments of 15

dat6 %>%
mutate(N2 = N + r2 * N * (1-N/K),
       N4 = N + r4 * N * (1-N/K),
       N10 =N + r10 * N * (1-N/K), 
       N20 = N + r20 * N * (1-N/K)) %>% 
ggplot(., aes(N, N2))+geom_line()+geom_line(aes(N,N), color='red', lty = 4)+
   geom_line(aes(N, N4), color='blue')+
   geom_line(aes(N, N10), color='green')+
   geom_line(aes(N, N20), color='orange')
             
r200 <- logistic_growth(2, 1000, 1, 100)
r20402 <- rbind(r20401, r200) 
r20402 %>% 
   ggplot(aes(year, N, color=r))+geom_line()
```

Look at the abundance for years 8-20. Is the population climbing steadily to its carrying capacity? What is it doing?

```{r}
r20402 %>% 
   ggplot(aes(year, N, color=r))+geom_line()+xlim(0, 20)
```

Try r = 2.1. Does the population reach carrying capacity? Will it ever? 
How about when r = 2.5? Does this look like any natural population you’ve heard about?

```{r}
dat7 <- data.frame(N = seq(0,1500,by=15), r2 = 0.2, r4 = 0.4, r10 = 1,r20 = 2, r21 = 2.1, r25 = 2.5, K= 1000) 

dat7 %>%
mutate(N2 = N + r2 * N * (1-N/K),
       N4 = N + r4 * N * (1-N/K),
       N10 =N + r10 * N * (1-N/K), 
       N20 = N + r20 * N * (1-N/K),
       N21 = N + r21 * N * (1-N/K),
       N25 = N + r25 * N * (1-N/K)) %>% 
ggplot(., aes(N, N2))+geom_line()+geom_line(aes(N,N), color='red', lty = 4)+
   geom_line(aes(N, N4), color='blue')+
   geom_line(aes(N, N10), color='green')+
   geom_line(aes(N, N20), color='orange')+
   geom_line(aes(N, N21), color='yellow')+
   geom_line(aes(N, N25), color=6)
             
r210 <- logistic_growth(2.1, 1000, 1, 100)
r250 <- logistic_growth(2.5, 1000, 1, 100)

r204025 <- rbind_all(list(r20402, r210, r250)) 
r204025 %>% 
   ggplot(aes(year, N, color=r))+geom_line()
```

Construct a bifurcation graph

First construct a dataframe of the values you want
```{r}
newd <- data.frame(r = seq(1.8,3, by=0.05), K = 1000, N = 1, t = 100)
```

run your function on the dataframe
```{r}
stored_array <- sapply(1:nrow(newd), function(i) logistic_growth(newd[i,1], newd[i,2],newd[i,3],newd[i,4]) )
```

Pull the array apart and create a dataframe
```{r}
a <- data.frame(stored_array[,1])

for(i in 2:nrow(newd)){
a <- rbind(a,data.frame(stored_array[,i]))
}
```

Remove all but the last ten years and plot results
```{r}
a %>% 
   filter(year>89) %>% 
   ggplot(aes(r, N, color = factor(r)))+geom_point()
```